Active deformation of the Shargyn Basin, a transpressional strike-slip intersection in western Mongolia.

### 2013-05-07 writing/paper/abstract.tex
writing.. writing.. writing..

\begin{abstractpage}
Intraplate faulting in central Asia is a major component of the Indo-Eurasian collision. The kinematics and mechanisms of intraplate deformation are important in understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources. I examine the fault kinematics surrouding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system. The basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many terminal thrust splays. The fault-kinematics, pre-existing structural weaknesses and a significant gravity anomaly all suggest that the basin basement is mechanically strong. This research demonstrates some of the potential kinematics for intraplate transpressional orogenies and emphasizes the importance of pre-existing crustal structure in the development of active faults.
\end{abstractpage}

>>>>

\begin{abstractpage}
Intraplate faulting in central Asia is a major component of the Indo-Eurasian collision. The kinematics and mechanisms of intraplate deformation are important in understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources. We examine the fault kinematics surrounding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system. The studies were performed using satellite data and targeted field transects. The results suggest the Shargyn basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many terminal thrust splays. The fault-kinematics, pre-existing structural weaknesses and a significant gravity anomaly all suggest that the basin basement is mechanically strong. This research demonstrates some of the potential kinematics for intraplate transpressional orogenies and emphasizes the importance of pre-existing crustal structure in the development of active faults.
\end{abstractpage}

### 2013-05-07 writing/paper/background.tex
writing.. writing.. writing..

\begin{figure}[h!]
	\centering
	\includegraphics[width = 6.5in, height = 4in]{broad.png}
	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes. Important locations or faults are labeled. The study area and location of Figure \ref{regional} is outlined in black. The difference between right-lateral transpressional strike-slip in the Altai Range and left-lateral transpressional strike-slip in the Gobi-Altai is especially important. Modified after \citet{Calais2003}}
	\label{broad}
\end{figure}

	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}. The region is dominated by NW-SE striking right-lateral slip. The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults. Others are linked by a right-slip faults. In some places, the thrusts are part of a restraining bend. Major faults, GPS motions, and earthquake moment tensors are shown in Figure \ref{broad}. Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.

	Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}. North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures. The deformation is diffuse over a 250-350km north-south extent. The range accomodate the eastwards and northeastwards deformation as measured by GPS. At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System \citep{Cunningham2003a}. There are no slip estimates for the Bogd Fault System in the north. However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.  Earthquake moment tensors show mainly left-lateral strike-slip in the Gobi-Altai.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width = 6in, height = 4in]{regional.png}
	\caption{Shaded relief map showing notable features in the Shargyn Basin region. Many of these faults are studied in this paper. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for the direction of slip.}
	\label{regional}
\end{figure}

>>>>

\begin{figure}[h!]
	\centering
	\includegraphics[width = 6.5in, height = 4in]{broad.png}
	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes. Important locations or faults are labeled. HTF = Hoh-Serh-Tsagaan-Salaa Fault, BF = Bolnay Fault, BGF = Bogd Fault. The study area and location of Figure \ref{regional} is outlined in black. The difference between right-lateral transpressional strike-slip in the Altai Range and left-lateral transpressional strike-slip in the Gobi-Altai is especially important. Modified after \citet{Calais2003}}
	\label{broad}
\end{figure}

	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}. The region is dominated by NW-SE striking right-lateral slip. The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults. Others are linked by a right-slip faults. In some places, the thrusts are part of a restraining bend. Major faults, GPS motions, and earthquake moment tensors are shown in Figure \ref{broad}. Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}. The geologic slip rate for a major transpressional fault, the Hoh-Serh-Tsagaan-Salaa Fault, is ~0.3 mm/yr of shortening and ~0.9 mm/yr of dextral slip. This consitutes ~20\% of the total strain accumulation in the Altai. \citep{Frankel2010}

	Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}. North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures. The deformation is diffuse over a 250-350km north-south extent. The range accomodate the eastwards and northeastwards deformation as measured by GPS. At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (shown as "GBTS" in Figure \ref{asia}) \citep{Cunningham2003a}. There are no slip estimates for the Bogd Fault System in the north. However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.  Earthquake moment tensors show mainly left-lateral strike-slip in the Gobi-Altai.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width = 6in, height = 4in]{intro.png}
	\caption{Shaded relief map showing notable features in the Shargyn Basin region. Many of these faults are studied in this paper. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for the direction of slip.}
	\label{regional}
\end{figure}
